Analysing Constraint Grammars with a SAT-solver

### 2016-02-24 paper/lrec/problem_desc.tex
add text in intro about SAT

By design, CG is a shallow and robust formalism. Individual rules can be written to address any property, such as ``verb'', ``copula verb in first person singular'', or ``the word form \emph{sailor} preceded by \emph{drunken} anywhere in the sentence''. This makes it possible to treat very particular edge cases without touching the more general rule: we would simply write the narrow rule first (``if noun AND \emph{sailor}''), and use a general rule (``if noun'') as catch-all.

@@

\begin{itemize}
\item[]\begin{verbatim}SELECT Inf IF (-1 Para OR De) (0C V) ;
SELECT Inf IF (-1 Prep) (0C V) ;
SELECT Inf IF (-1C Vai) ;
SELECT Inf IF (-1C Vbmod) (0C V) ;
SELECT Inf IF (-1C Ter/de) ;
SELECT Inf IF (-1C Vbmod) (0 Ser) ;
\end{verbatim}
\end{itemize}

\caption{Rules to select infinitive.}
\end{center}

@@

While rule-internal conflicts can be detected by simpler means, taking
care of rule interaction requires a {\em semantic} rather than a {\em syntactic} analysis.
%In addition to analysing the whole grammar,
%Our method can be used to
%check individual rules. For instance,

>>>>

By design, CG is a shallow and robust formalism. 
There is no particular hierarchy between lexical, morphological,
syntactic or even semantic tags: individual rules can be written to address any
property, such as ``verb'', ``copula verb in first person singular'',
or ``the word form \emph{sailor}, preceded by \emph{drunken} anywhere in the
sentence''. This makes it possible to treat very particular edge
cases without touching the more general rule: we would simply write
the narrow rule first (``if noun AND \emph{sailor}''), and introduce
the general rule (``if noun'') later.

@@

\begin{itemize}
\item[]
\begin{verbatim}SELECT Inf IF (-1 Para OR De) (0C V) ;
SELECT Inf IF (-1 Prep) (0C V) ;
SELECT Inf IF (-1C Vai) ;
SELECT Inf IF (-1C Vbmod) (0C V) ;
SELECT Inf IF (-1C Ter/de) ;
SELECT Inf IF (-1C Vbmod) (0 Ser) ;
\end{verbatim}
\end{itemize}

\caption{Rules to select infinitive in Portuguese.}
\end{center}

@@

While rule-internal conflicts can be detected by simpler means, taking
care of rule interaction requires a {\em semantic} rather than a {\em
 syntactic} analysis.
In order to find effects of rule interaction, we must keep track of
the possible sentences at each step. After each rule, we have two
possibilities: the rule fires, or it does not fire. In case the rule does
not fire, we have again two options: either its conditions are not met,
or its target is the only remaining analysis.

We express these requirements as a \emph{Boolean satisfiability problem} (SAT).
SAT problems consist of two components: a set of Boolean variables, and a set
of clauses on those variables. For instance, let the set be $\{a,b\}$
and the formulas $\{a \vee b, \neg{}a\}$. A program called \emph{SAT
  solver} will try to find a solution, where all the variables have a
value. For this particular problem, the unique solution is $a=False,
b=True$. It is also possible for a SAT problem to have no solution, or
multiple solutions.

%these possibilities as SAT variables, and ask for a solution: ``after all
%rules in $R$ have applied, is there an input that will make $r'$ fire?'' 
%To our knowledge, there is no other CG tool that would keep track of
%previously executed rules
